Considering scale in multilayer animal social networks

### 2020-06-21 paper/manuscript.Rmd
order foraging, forest, open

## Landscape data and habitat classification
Land cover data were provided by the Newfoundland and Labrador Wildlife Division
[@Integrated_2013]<!--TODO: check ref-->. Available land cover classification
included nine habitat types at 30 m resolution. We reclassified the land cover
types into three categories using the R package `raster` [@Hijmans_2019]:
foraging, open, and forest habitats. Foraging habitat consisted of foraging, open
habitat consisted of wetland, rocky barrens, and anthropogenic, while forest
habitat consisted of coniferous forest, conifer scrub, broad leaf forest, and
mixed-wood forest. Water habitat was excluded from all subsequent analyses.

@@

We generated multilayer networks across a series of social distance thresholds 
for group assignment. Social distance thresholds (5, 25, 50, 75, 100, 250 and
500 m) represented a range of visual (short distance) to auditory (long
distance) sensory modalities of caribou. Multilayer networks consisted of the
social association between `r DT[, uniqueN(get(idcol))]` individuals across
three habitat layers (open, forest and foraging) for the entire study period. At
the finest scale, individuals within 5 m of one another were considered in the
same group, whereas at the coarsest scale, individuals within 500 m of one
another were considered in the same group. Increasing social distance threshold 
is a proxy for potential behavioural interactions that occur across these distances.
We calculated graph strength and
multidegree for each habitat layer and social distance threshold, and edge
overlap across the entire series of networks to determine proportion of total
edges observed in each combination of habitat layer and spatial distance
threshold.

### Spatial scale
To assess the influence of spatial scale, we aggregated the land cover raster
using the R package `grainchanger` [@Graham_2019] across a series of scales (100
- 1000 m by steps of 100 m). The land cover raster was aggregated using a modal
moving window method using a circular window corresponding to the above scales
[@Graham_2019a]. We assume 30 m and 100 m represents fine-scale decision making
for caribou during foraging, while re-sampling at 500 m , 750 m and 1000 m
represents the scale at which caribou tend to select and avoid habitat
(Bastille-Rousseau et al. 2017). For each spatial scale, we generated a
multilayer network consisting of the social associations between 
`r DT[, uniqueN(get(idcol))]` individuals across three habitat layers (open, forest and
foraging) for the entire study period. Social associations of individuals were
defined by spatiotemporal overlap within 5 minutes and 50 m. We calculated graph
strength and multidegree for each habitat layer and spatial scale, and edge
overlap across the entire series of networks to determine proportion of total
edges observed in each combination of habitat layer and spatial scale.
<!-- JWT: Your graphs present things in the order of Foraging, Forest, Open, but in the text you do open, forest, forage. They need to be consistent -->

@@

#### Number of observations
Studies of social network analysis vary in the number and frequency of
observations as well as the data collection technique used to generate networks
[@Davis_2018; @Webber_2019]. For example, GPS data is commonly collected at a
fixed rate, e.g. every $x$ minutes or hours, continuously throughout the study
period. Fix rate is a reflection of number of observations an individual would
have been observed and recorded in traditional ethological studies. To
investigate the influence of number of observations, we generated multilayer
networks of social association between `r DT[, uniqueN(get(idcol))]` individuals
across three habitat layers (open, forest and foraging) using $N$ randomly
selected observations. We first generated a maximum of 1000 timesteps and
iteratively used $N$ of these ($N$ = 10 - 1000 observations by steps of 10), to
ensure previously included timesteps, and resulting associations, were
accumulated with subsequent observations to mimic collection of observational
data. Within each multilayer network with $N$ observations, we calculated
individual graph strength and multidegree. We calculated edge overlap across the
entire series of networks to determine proportion of total edges as number of
observations increased.

# Results
Individuals in multilayer networks became more connected as the spatial distance
threshold increased. Mean graph strength differed across habitat layers with
similar and higher graph strength in open and foraging layers, and lower strength
in forest layers (Figure \@ref(fig:socres)). Individual graph strength increased
sharply between 5 m and 100 m in all three habitat classes followed by a plateau in
the rate of increase after 100 m (<!--TODO: figure XX-->). 
These results suggest the optimal social scale at
which groups should be assigned is between ~20 - 100 m based on
the rate at which network metrics <!--QW: which metrics?--> stabilized. For ungulates, groups are often
assigned based on 50 m threshold (Lingle 2003; Lesmeries et al. 2018; Peignier
et al. 2019), suggesting that given similar quantities of data, ~50 m is likely
an appropriate social scale for generating social networks in caribou and
similar species. <!--QW: I'm not sure about this statement. Should it be more about 
how GPS data allows (e.g. from this study and Peignier et al; Lesmerises et al.) us to confirm 
the use of 50m buffer that has long been used for studies in the field (e.g. Lingle 2003 and 
any Clutton-Brock red deer paper that uses group size data)-->

Increasing land cover resolution resulted in decreased availability of foraging habitat,
a rare and patchy habitat, and corresponding low connectivity of individuals in
these habitats at higher resolution (Figure \@ref(fig:lcres)). Graph strength
and edge overlap were relatively consistent across land cover resolutions for
forest and open layers (Figure \@ref(fig:lcres) D, E). Edge overlap decreased
sharply between 30 m and 600 m in foraging layers, afterwards remaining stable to
1000 m (Figure \@ref(fig:lcres) D). Graph strength was more variable within
foraging habitat layers across land cover resolutions than open and forest habitat layers (Figure
\@ref(fig:lcres) D). <!-- JWT: the order things are discussed here is a little confusing, you discuss strength, edge overlap, then strenght again. I think this should be reordered so it's easier to follow. -->
The proportion of relocations in forest and open habitats
increased with increasing spatial resolution, while decreasing in foraging habitat
(Supplementary 1 Figure 1).  Overall, these results indicate the importance of
matching land cover resolution to scale of selection as well as ensuring the
land cover product used has a resolution fine enough to detect landscape
features and habitats relevant to study species, such as important patchy and
rare habitats.

>>>>

## Landscape data and habitat classification
Land cover data were provided by the Newfoundland and Labrador Wildlife Division
[@Integrated_2013]<!--TODO: check ref-->. Available land cover classification
included nine habitat types at 30 m resolution. We reclassified the land cover
types into three categories using the R package `raster` [@Hijmans_2019]:
foraging, forest, and open habitats. Foraging habitat consisted of foraging,
forest habitat consisted of coniferous forest, conifer scrub, broad leaf forest,
and mixed-wood forest, while open habitat consisted of wetland, rocky barrens,
and anthropogenic. Water habitat was excluded from all subsequent analyses.

@@

We generated multilayer networks across a series of social distance thresholds 
for group assignment. Social distance thresholds (5, 25, 50, 75, 100, 250 and
500 m) represented a range of visual (short distance) to auditory (long
distance) sensory modalities of caribou. Multilayer networks consisted of the
social association between `r DT[, uniqueN(get(idcol))]` individuals across
three habitat layers (foraging, forest and open) for the entire study period. At
the finest scale, individuals within 5 m of one another were considered in the
same group, whereas at the coarsest scale, individuals within 500 m of one
another were considered in the same group. Increasing social distance threshold 
is a proxy for potential behavioural interactions that occur across these distances.
We calculated graph strength and
multidegree for each habitat layer and social distance threshold, and edge
overlap across the entire series of networks to determine proportion of total
edges observed in each combination of habitat layer and spatial distance
threshold.

### Spatial scale
To assess the influence of spatial scale, we aggregated the land cover raster
using the R package `grainchanger` [@Graham_2019] across a series of scales (100
- 1000 m by steps of 100 m). The land cover raster was aggregated using a modal
moving window method using a circular window corresponding to the above scales
[@Graham_2019a]. We assume 30 m and 100 m represents fine-scale decision making
for caribou during foraging, while re-sampling at 500 m , 750 m and 1000 m
represents the scale at which caribou tend to select and avoid habitat
(Bastille-Rousseau et al. 2017). For each spatial scale, we generated a
multilayer network consisting of the social associations between 
`r DT[, uniqueN(get(idcol))]` individuals across three habitat layers (foraging, forest and
open) for the entire study period. Social associations of individuals were
defined by spatiotemporal overlap within 5 minutes and 50 m. We calculated graph
strength and multidegree for each habitat layer and spatial scale, and edge
overlap across the entire series of networks to determine proportion of total
edges observed in each combination of habitat layer and spatial scale.
<!-- JWT: Your graphs present things in the order of Foraging, Forest, Open, but in the text you do open, forest, forage. They need to be consistent -->

@@

#### Number of observations
Studies of social network analysis vary in the number and frequency of
observations as well as the data collection technique used to generate networks
[@Davis_2018; @Webber_2019]. For example, GPS data is commonly collected at a
fixed rate, e.g. every $x$ minutes or hours, continuously throughout the study
period. Fix rate is a reflection of number of observations an individual would
have been observed and recorded in traditional ethological studies. To
investigate the influence of number of observations, we generated multilayer
networks of social association between `r DT[, uniqueN(get(idcol))]` individuals
across three habitat layers (foraging, forest and open) using $N$ randomly
selected observations. We first generated a maximum of 1000 timesteps and
iteratively used $N$ of these ($N$ = 10 - 1000 observations by steps of 10), to
ensure previously included timesteps, and resulting associations, were
accumulated with subsequent observations to mimic collection of observational
data. Within each multilayer network with $N$ observations, we calculated
individual graph strength and multidegree. We calculated edge overlap across the
entire series of networks to determine proportion of total edges as number of
observations increased.

# Results
Individuals in multilayer networks became more connected as the spatial distance
threshold increased. Mean graph strength differed across habitat layers with
similar and higher graph strength in foraging and open layers, and lower strength
in forest layers (Figure \@ref(fig:socres)). Individual graph strength increased
sharply between 5 m and 100 m in all three habitat classes followed by a plateau in
the rate of increase after 100 m (<!--TODO: figure XX-->). 
These results suggest the optimal social scale at
which groups should be assigned is between ~20 - 100 m based on
the rate at which network metrics <!--QW: which metrics?--> stabilized. For ungulates, groups are often
assigned based on 50 m threshold (Lingle 2003; Lesmeries et al. 2018; Peignier
et al. 2019), suggesting that given similar quantities of data, ~50 m is likely
an appropriate social scale for generating social networks in caribou and
similar species. <!--QW: I'm not sure about this statement. Should it be more about 
how GPS data allows (e.g. from this study and Peignier et al; Lesmerises et al.) us to confirm 
the use of 50m buffer that has long been used for studies in the field (e.g. Lingle 2003 and 
any Clutton-Brock red deer paper that uses group size data)-->

Increasing land cover resolution resulted in decreased availability of foraging habitat,
a rare habitat, and corresponding low connectivity of individuals in
these habitats at higher resolution (Figure \@ref(fig:lcres)). Graph strength
and edge overlap were relatively consistent across land cover resolutions for
forest and open layers (Figure \@ref(fig:lcres) D, E). Edge overlap decreased
sharply between 30 m and 600 m in foraging layers, afterwards remaining stable to
1000 m (Figure \@ref(fig:lcres) D). Graph strength was more variable within
foraging habitat layers across land cover resolutions than open and forest habitat layers (Figure
\@ref(fig:lcres) D). <!-- JWT: the order things are discussed here is a little confusing, you discuss strength, edge overlap, then strenght again. I think this should be reordered so it's easier to follow. -->
The proportion of relocations in forest and open habitats
increased with increasing spatial resolution, while decreasing in foraging habitat
(Supplementary 1 Figure 1).  Overall, these results indicate the importance of
matching land cover resolution to scale of selection as well as ensuring the
land cover product used has a resolution fine enough to detect landscape
features and habitats relevant to study species, such as important patchy and
rare habitats.
